---
title: "Lab4_Demo"
author: "Mateo Robbins"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting
library(rsample)   # for data splitting
library(recipes)   # data preprocessing
library(caret)     # for logistic regression modeling
```
Let's explore how employee income and overtime hours worked affect likelihood of employee attrition.  Any predictions?

```{r}
#
data("attrition", package = "modeldata")

df <- attrition %>% mutate_if(is.ordered, factor, ordered = FALSE)

# Create training (70%) and test (30%) sets for the 
# rsample::attrition data.
set.seed(123)  # for reproducibility 
churn_split <- initial_split(df, prop = .7)
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```

Let's set up set up our recipes now for preprocessing. 
```{r recipe}
#specify and prep recipe
churn_rec <- recipe(Attrition ~ ., data = churn_train) %>% 
  step_integer(Attrition, zero_based = TRUE) %>% 
  prep(churn_train)

#bake recipe with training data
churn_baked_train <- bake(churn_rec, new_data = churn_train)
```

```{r specify_models_glm}
#MonthlyIncome
model_inc <- glm(data = churn_baked_train, Attrition ~ MonthlyIncome, family = "binomial")
  
#OverTime
model_time <- glm(data = churn_baked_train, Attrition ~ OverTime, family = "binomial")
```


```{r tidy_model_objs}
broom::tidy(model_inc)
broom::tidy(model_time)
```

```{r exp_coefs}
#exponentiate the coefficients from model objects for interpretation. Gives us changes in odds of attrition
exp(coef(model_inc))
exp(coef(model_time))
```
The odds of an employee leaving increases multiplicatively (aka mulitplied) by 0.99 by every one additional dollar per month (therefore likelihood of them leaving is incrementally decreasing)

The odds of an employee leaving mulitply by 3.5 times if they worked overtime.

```{r recode_attrition_test}
churn_baked_test <- bake(churn_rec, new_data = churn_test)
```

```{r plot_income_attrition}
ggplot(churn_baked_test, aes(x = MonthlyIncome, y = Attrition)) +
  geom_point() +
  stat_smooth(method = "glm", se = TRUE, method.args = list(family = binomial))
```

We can add more predictors, creating a multiple logistic regression model

```{r mult_log_regression}
model_both <- glm(Attrition ~ MonthlyIncome + OverTime, family = "binomial", data = churn_train)

broom::tidy(model_both)
```
```{r}
ggplot(churn_baked_test, aes(x = MonthlyIncome, y = Attrition, color = OverTime)) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "glm", se = FALSE, method.args = list(family = binomial))
```

